Reviewer comments from PeerJ

### 2015-03-25 PeerJResponse.tex
final edits revision

\lev{Okay. I am not sure that we have significance values for these.}

\mbh{We don't and it's kind of a strange/silly request...Jeff and I will deal with this}
\jri{ i can do the simple but dumb calculation if you guys want for response to reviewers, but disagree we should include in manuscript. we mention they are at the tail of the empirical distribution from wright et al don't we?}
\mbh{I agree this is a silly request and that we shouldn't include in the manuscript; yes, we indicate Taj.D for our pops is in the extreme tail based on comparison to both Wright et al. and Tiffin and Moeller}

@@

\lev{Okay. What do you guys think of this? I don't think we have tests of significance for tajd}

\mbh{Jeff and I will address this}
\jri{sloppy my ass. <censored> reviewer!}
\lev{ :) }

>>>>

\res{We have determined significance of the Tajima's D results by comparing our empirical values to coalescent simulations generated using the program ms \citep{Hudson2002}.  We describe these methods and results on page 7, lines 157-161 and page 13, lines 368-369.  Significant values are now marked with an asterisk in Table 2}

@@

\res{Please see our response to item 14.}

### 2015-03-25 Vann_et_al._Manuscript.tex
final edits revision

Maize (\emph{Zea mays} ssp. \emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 9,000 B.P. in southwest Mexico \citep{Piperno2009, Matsuoka2002}. Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}. 
\track{Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, with \emph{parviglumis} distributed in the lowlands of Mexico and \emph{mexicana} in the highlands. Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}}.

@@

\textbf{Sampling and genotyping---}\track{We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations. Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites.} DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb). \track{Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig. S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article). Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is located. Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR. Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2, See Supplemental Materials with the online version of this article).} Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}. These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled. To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between \emph{Hopscotch} frequency and environmental variables using the software BayEnv \citep{Coop2010}. BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}). We used teosinte (ssp. \emph{parviglumis} and ssp. \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker. \track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}. Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples. Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.

@@

\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the Analysis package from the  Libsequence library \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D). To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (10,000 repetitions). For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.

@@

\textbf{Genotyping for the \emph{Hopscotch} insertion---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened \track{(Table S1 and Table S2, See Supplemental Materials with the online version of this article).}Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig2Map}). Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) (Table \ref{Table1Fst}) are similar to genome-wide estimates from previous studies \citealt{Pyhajarvi2013}. Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

\textbf{Sequencing upstream regions of the \emph{tb1} ORF---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations \track{from the Jalisco region}. After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively. For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values. Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}). For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}). Tajima's D is positive \track{in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants} (Table \ref{Table2Diversity}). Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).

\textbf{Evidence of introgression around the \emph{tb1} region---} \track{We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}.} The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize. Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte. To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome 1 and specifically in the \emph{tb1} region. If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element. We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.\ref{Fig3LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion. In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).

@@

Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings. Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance. A bottleneck in San Lorenzo, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster. Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations. 
Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in populations of \emph{parviglumis}. Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative across populations (Table \ref{Table2Diversity}). In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}. Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.  Finally, while these results are consistent with selection at the \emph{tb1} locus in teosinte, they do not confirm selection specifically on the \emph{Hopscotch} insertion at this locus.

Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}. \citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}. Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte. We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize. \track{It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}. This factor may have limited our capacity to observe variation in tillering index.}

@@

% Table 2 Diversity Statistics
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering
  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
    \begin{tabular}{ccccc}\\\toprule
    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
    \end{tabular}
  \label{Table2Diversity}
\end{table}

>>>>

Maize (\emph{Zea mays} ssp. \emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 9,000 B.P. in southwest Mexico \citep{Piperno2009, Matsuoka2002}. Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}. 
Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, with \emph{parviglumis} distributed in the lowlands of Mexico and \emph{mexicana} in the highlands. Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}.

@@

\textbf{Sampling and genotyping---}We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations. Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites. DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb). Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig. \ref{Fig1Cartoon} and Fig. S1; See Supplemental Materials with the online version of this article). Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is located. Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR. Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2; See Supplemental Materials with the online version of this article). Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}. These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled. To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between \emph{Hopscotch} frequency and environmental variables using the software BayEnv \citep{Coop2010}. BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}). We used teosinte (ssp. \emph{parviglumis} and ssp. \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker. SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}. Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples. Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.

@@

\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the Analysis package from the  Libsequence library \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D). Significance of Tajima's D results was gauged by comparing empirical data to 10,000 coalescent simulations conducted using the program ms \citep{Hudson2002} under a standard neutral model based on observed estimates of the population mutation rate theta and assuming an identical value for the population recombination rate rho. Empirical results falling outside the 95\% confidence interval of our simulated data were deemed significant. To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (10,000 repetitions). For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.

@@

\textbf{Genotyping for the \emph{Hopscotch} insertion---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article). Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig. \ref{Fig2Map}). Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes. We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) (Table \ref{Table1Fst}) are similar to genome-wide estimates from previous studies \citealt{Pyhajarvi2013}. Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

\textbf{Sequencing upstream regions of the \emph{tb1} ORF---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region. After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively. For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values. Additionally, Tajima's D is significantly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}). For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}). Tajima's D is positive in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants (Table \ref{Table2Diversity}). Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).

\textbf{Evidence of introgression around the \emph{tb1} region---} We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citet{Pyhajarvi2013}. The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize. Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte. To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome 1 and specifically in the \emph{tb1} region. If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element. We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.\ref{Fig3LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion. In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).

@@

Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings. Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance. A bottleneck in San Lorenzo, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster. Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations. 
Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in populations of \emph{parviglumis}. Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative across populations (Table \ref{Table2Diversity}). In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007} and significantly negative in three of our surveyed populations (EjuA, EjuB, MSA) based on coalescent simulations under a standard neutral model. Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.  Finally, while these results are consistent with selection at the \emph{tb1} locus in teosinte, they do not confirm selection specifically on the \emph{Hopscotch} insertion at this locus.

Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}. \citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}. Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte. We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize. It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}. This factor may have limited our capacity to observe variation in tillering index.

@@

% Table 2 Diversity Statistics
\begin{table}[htbp]
\rowcolors{2}{gray!25}{white}
  \centering
  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus. Significant values are marked with an asterisk}.
    \begin{tabular}{ccccc}\\\toprule
    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650* \\
    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831* \\
    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755* \\
    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
    \end{tabular}
  \label{Table2Diversity}
\end{table}
